---
title: "taganalysis_text2vec"
author: "Sandy HE"
date: "5 April 2018"
output: html_document
---

```{r setup, include=FALSE}
library(text2vec)  
library(data.table) 
library(stringr)
```

# Preliminary 
## Initialization
* Collect tags data  
 In the input file, tag info like this: happy\*100;indie rock\*82;rock\*65
* Set stopwords for removal
* Divide track id for train and test set

```{r initial, include=FALSE}
#initial
alltags <- fread("./statwords.csv")
#use top40 frequent words calculated in the experiment before
tematrix <- fread("tagsemomatrix.csv")
topwords <- tematrix$V1

#divide into train set and test set with the proportion of 4:1
all_ids <- alltags$track_id 
train_ids <- sample(all_ids, nrow(alltags)/5*4)
test_ids <- setdiff(all_ids, train_ids) 
```

The function of tags transform
For one song, merge all of tags into one string with the format like this:
[tag1]\*popularity1 + [tag2]\*popularity2+...+[tagN]\*popularityN
* Input: tags info of each song
* Output: merged tags info
```{r itemsmultitrans}
itemsMultiTrans <- function(items){
  itemlist <- tstrsplit(items, "\\*")
  itemdf <- as.data.frame(itemlist, stringsAsFactors = F,col.names = c("item","popularity"))
  if(!is.integer(itemdf$popularity)) print(itemdf)
  itemdf$multiItem <- strrep(paste0(itemdf$item," "),as.integer(itemdf$popularity) )
  itemdf <- itemdf[itemdf$popularity!=0,]
  
  itemsstr<-paste(itemdf$multiItem,collapse =" ")
  itemsstr
}
```

## Process Tags and Merge into String 
lowercase-> split-> tags multipled with popularity and merge
```{r alltags, eval = FALSE}
obsv <- str_to_lower(alltags$tags)
obsv <- str_split(obsv,pattern = ";")      #list
obsv1 <- sapply(obsv, itemsMultiTrans)    #char vector

#data backup
alltags$transtags <- obsv1
#fwrite(alltags,file = "alltranstags.csv")
```

```{r alltagstable}
alltags <- fread("alltranstags.csv")
knitr::kable(
  alltags[1:5,c(1,3)],
  caption = "transformed tags info"
)
```

## prepare train set and test set
```{r dataset}
train = alltags[alltags$track_id %in% train_ids,]  
test = alltags[alltags$track_id %in% test_ids,]  
```

## tags analysis
* generate token for training dataset
* create vocabulary with stop words removal
* remove low-freq and high-freq words in vocabulary
* generate word vector
```{r wordvec}
tok_fun <- word_tokenizer  
it_train <- itoken(train$transtags,   
                  #preprocessor = prep_fun,   
                  tokenizer = tok_fun,   
                  ids = train$track_id,   
                  progressbar = F)  

#set stop words and remove them from vocabulary
stop_words <- c("the","to","of","and","a","in","this","-","for","it","that","on","is")
system.time(vocab <- create_vocabulary(it_train,stopwords = stop_words))

#remove low-freq and high-freq words
pruned_vocab = prune_vocabulary(vocab,   
                                term_count_min = 10,   
                                doc_proportion_max = 0.5,  
                                doc_proportion_min = 0.001)  

#set up corpus vector
vectorizer <- vocab_vectorizer(pruned_vocab)
```


## Generate DTM(doc-term matrix). 
Here we can understand it as song-term matrix
```{r DTM, eval=FALSE}
system.time(dtm_train <- create_dtm(it_train, vectorizer))
```

## Generate TCM(term-co-occurrence matrix).
```{r TCM}
system.time(tcm_train <- create_tcm(it_train, vectorizer, skip_grams_window = 5L,
           skip_grams_window_context = c("symmetric", "right", "left")))
```

## Prune Sparse Matrix Containing topN-freq Tags 
```{r pruned_tcm}
tcm_sample <- tcm_train[rownames(tcm_train) %in% topwords,colnames(tcm_train) %in% topwords]
tcm_sample_mat <- as.matrix(tcm_sample)

tcm_sample_diag <- diag(diag(tcm_sample_mat))
tcm_sample_whole <- tcm_sample_mat + t(tcm_sample_mat) - tcm_sample_diag
tcm_sample_whole <- round(tcm_sample_whole,2)
tcm_sample_df <- as.data.frame(tcm_sample_whole)

#tags <- rownames(tcm_sample_df)
#tcm_sample_df<- cbind(tags,tcm_sample_df)
#fwrite(tcm_sample_df, file = "tcm_sample_pruned_vocab.csv")
knitr::kable(
  tcm_sample_df,
  caption = "tags co-occurrence matrix"
)
```
