---
title: "How to map emotion tags to dimensional space"
output:
  html_document: default
  word_document: default
date: '2018-03-12'
---


```{r setup, include = FALSE}
library(stringr)
library(data.table)
library(ggplot2)
library(dplyr)
```
# Preliminary 
## Initialization
* Collect tags data  
 In the input file, tag info like this: happy\*100;indie rock\*82;rock\*65
* Set stopwords for removal
* Set genre words for removal
```{r alltags, include = FALSE}
#initial
alltags <- fread("./statwords.csv")
#in the final statistical form, remove some meaningless words
exword <- c("the","to","of","and","a","in","this","-","for","it","that","on","is")
#exgenre <- c("pop","rock","punk","jazz","folk","indie","classic","electronic","latin")
exgenre <- c("pop|rock|punk|jazz|folk|indie|classic|electronic|latin|hip hop|blues|rnb|rap|vocal|country|ballad")
```

## This function is extracting emotion tags info and tag frequency and weight.
* Input: ekey - specify emotion type
* Output: one list contains two data frames. 
* One dataset is single words statistic. The other is phrases statistic.
```{r statwords(), eval = FALSE}
statwords <- function(ekey){
  result <- grep(ekey,alltags$tags,ignore.case = TRUE)
  if(length(result)==0 ){
    return(NULL)
  }
  subtags <- alltags[result,]
  obsv <- str_to_lower(subtags$tags)
  obsv <- str_split(obsv,pattern = ";")      #list
  obsv1 <- unlist(obsv)                      #phrases vector with '*' and weight
  obsv2 <- sapply(strsplit(obsv1, "\\*"), "[", 1)  #phrases vector
  obsv3 <- sapply(strsplit(obsv1, "\\*"), "[", 2)  #weight vector
  
  
  #keep original phrases
  obsv4 <- as.data.frame(obsv2, stringsAsFactors = FALSE)   #data frame
  names(obsv4) <- c("phrases")
  obsv4$freq <- 1L
  obsv4$weight <- as.numeric(obsv3)
  #obsv4 <- obsv4[!(obsv4$phrases %in% exgenre),] 
  obsv4 <- obsv4[!(grepl(exgenre,obsv4$phrases)),] 
  
  #split into single words
  obsv5 <- paste0(obsv2,collapse=" ")  #only one string containing all words
  obsv6 <- str_split(obsv5,pattern = " ") 
  obsv6 <- unlist(obsv6)               # word vector
  
  
  #split single weight for each word
  times <- lengths(str_split(obsv2,pattern = " "))
  obsv7 <- as.data.frame(obsv6, stringsAsFactors = FALSE)   #data frame
  names(obsv7) <- c("words")
  obsv7$freq <- 1L
  obsv7$weight <- as.numeric(rep(obsv3,times))
  obsv7 <- obsv7[!(obsv7$words %in% exword),]  
  obsv7 <- obsv7[!(obsv7$words %in% exgenre),]  
  
  #add obsv4,obsv7 into one list as return
  list(obsv4,obsv7)
  
}
```

## Generate statistic data
* Setting keyword, then use function statwords(key) to generate tag frequency and weight for specified emotion type
* For each emotion type, calculate the statistic tags information
* Here we set key word to **happy**, **sad**, **angry**, **relax**, **calm** and run this part of code for them separately
* Export statistic data as csv files
```{r statwordswithemotion, eval = FALSE}

keyword <- c("happy")
statlist <- statwords(keyword)
if(!is.null(statlist)){
  form1 <- statlist[[1]]
  form2 <- statlist[[2]]
}

#==== data handling ====
# phrase statistics with frequency
statwords1 <- aggregate(freq~phrases,data=form1,FUN=sum)
statwords1 <- statwords1[order(statwords1$freq, decreasing = TRUE),]
rownames(statwords1) <- c()
fwrite(statwords1,file = paste0(keyword,"_statwords1.csv") )

# phrase statistics with weight*frequency
statwords2 <- aggregate(weight~phrases,data=form1,FUN=sum)
statwords2 <- statwords2[order(statwords2$weight, decreasing = TRUE),]
rownames(statwords2) <- c()
fwrite(statwords2,file = paste0(keyword,"_statwords2.csv") )
```

# Mapping method
## Initialize emotion statistic data
```{r emotionstatwords}
sadwords <- as.data.frame(fread("sad_statwords2.csv"))
happywords <- as.data.frame(fread("happy_statwords2.csv"))
angrywords <- as.data.frame(fread("angry_statwords2.csv"))
relaxwords <- as.data.frame(fread("relax_statwords2.csv"))
calmwords <- as.data.frame(fread("calm_statwords2.csv"))
knitr::kable(
  happywords[1:5, ], 
  caption = "Take happy words dataset as an example. It looks like this:"
)
```


## Get top40-weight value for all emotion types
1. Get top80-weight tags for each emotion type
2. Merge them to get unique top-weight tags
3. Calculate *commonwords* for all emotion types
4. Keep top30-weight tags for each emotion type, then merge and remove duplicated tags to get *uniquewords*
5. Remove *commonwords* from *uniquewords*   
*[**TBD**]*I believe R has package to do this better than me.It's about NLP and Emotion Analysis. I will try later. 
```{r commonwords}
top <- 80L
topsadwords <- head(sadwords,top)
tophappywords <- head(happywords,top)
topangrywords <- head(angrywords,top)
toprelaxwords <- head(relaxwords,top)
#topcalmwords <- head(calmwords,top)
  
#totalwords <- rbind(topsadwords,tophappywords,topangrywords,topcalmwords)
totalwords <- rbind(topsadwords,tophappywords,topangrywords,toprelaxwords)
totalwords$freq <- 1L

uniquewords <- aggregate(freq~phrases,data=totalwords,FUN=sum)
uniquewords <- uniquewords[order(uniquewords$freq, decreasing = TRUE),]
rownames(uniquewords) <- c()


commonwords <- uniquewords[uniquewords$freq==4,]$phrases
print(commonwords)
```
Totally, there are `r length(commonwords)` common words.   

```{r uniquewords }
top <- 30L
top30sadwords <- head(sadwords,top)
top30happywords <- head(happywords,top)
top30angrywords <- head(angrywords,top)
top30relaxwords <- head(relaxwords,top)
top30calmwords <- head(calmwords,top)
#totaltop30words <- rbind(top30sadwords,top30happywords,top30angrywords,top30calmwords)
totaltop30words <- rbind(top30sadwords,top30happywords,top30angrywords,top30relaxwords)
totaltop30words$freq <- 1L

uniquewords <- aggregate(freq~phrases,data=totaltop30words,FUN=sum)
uniquewords <- uniquewords[order(uniquewords$freq, decreasing = TRUE),]
rownames(uniquewords) <- c()

uniquewords <- uniquewords[!(uniquewords$phrases %in% commonwords),]$phrases
print(uniquewords)
```
Finally, `r length(uniquewords)` top-weight tags is selected.   

##Calculate emotion-tags matrix
```{r etmatrix1}
uniquehappywords <- happywords[happywords$phrases %in% uniquewords,]
colnames(uniquehappywords) <- c("phrases","happy")
uniquesadwords <- sadwords[sadwords$phrases %in% uniquewords,]
colnames(uniquesadwords) <- c("phrases","sad")
uniqueangrywords <- angrywords[angrywords$phrases %in% uniquewords,]
colnames(uniqueangrywords) <- c("phrases","angry")
uniquerelaxwords <- relaxwords[relaxwords$phrases %in% uniquewords,]
colnames(uniquerelaxwords) <- c("phrases","relax")
uniquecalmwords <- calmwords[calmwords$phrases %in% uniquewords,]
colnames(uniquecalmwords) <- c("phrases","calm")

etmatrix <- merge(uniquehappywords,uniquesadwords,by="phrases")
etmatrix <- merge(etmatrix,uniqueangrywords,by="phrases")
#etmatrix <- merge(etmatrix,uniquecalmwords,by="phrases")
etmatrix <- merge(etmatrix,uniquerelaxwords,by="phrases")
row.names(etmatrix) <- etmatrix$phrases

#==== tags-emotion matrix====
etmatrix1 <- as.matrix(etmatrix[,2:5])
knitr::kable(
  etmatrix1[1:10,],
  caption = "tags-emotion matrice"
)
```

The tags-emotion matrix size is `r nrow(etmatrix1)`*`r ncol(etmatrix1)`   

## Balancing tags data size for each emotion type
### Prepare songs tag data
```{r songstag, eval=FALSE}
#==== build songs set covering 4 emotion types====
alltags <- fread("./statwords.csv")
result <- grep(keyword,alltags$tags,ignore.case = TRUE)
subtags <- alltags[result,]
fwrite(subtags,paste0("tagswith",keyword,".csv"))
```

### Balance data
```{r songstags, include=FALSE}
happydata <- as.data.frame(fread("tagswithhappy.csv"))
saddata <- as.data.frame(fread("tagswithsad.csv"))
angrydata <- as.data.frame(fread("tagswithangry.csv"))
relaxdata <- as.data.frame(fread("tagswithrelax.csv"))
calmdata <- as.data.frame(fread("tagswithcalm.csv"))


samp <- sample(nrow(happydata), nrow(angrydata)) 
samp <- sort(samp)
happydata <- happydata[samp,]

samp <- sample(nrow(saddata), nrow(angrydata))
samp <- sort(samp)
saddata <- saddata[samp,]

samp <- sample(nrow(relaxdata), nrow(angrydata))
samp <- sort(samp)
relaxdata <- relaxdata[samp,]

samp <- sample(nrow(calmdata), nrow(angrydata))
samp <- sort(samp)
calmdata <- calmdata[samp,]

songstags <- rbind(happydata,saddata,angrydata,relaxdata)
#songstags <- rbind(happydata,saddata,angrydata,calmdata)
songstags <- unique(songstags$track_id)
songstags <- alltags[alltags$track_id %in% songstags,]
```
Use **angry** data size `r nrow(angrydata)` as the base size, other emotion data is sampled. 
```{r}
knitr::kable(
  head(songstags),
  caption = "Balanced songs tag dataset"
)
```

## Calculate songs-tags matrix
Here is the function that calculate the weight of topN unique words for each song track
```{r strtrans()}
strtrans <- function(x){
  obsv1 <- strsplit(x,"\\*")
  obsv2 <- data.frame(matrix(unlist(obsv1),ncol = 2, byrow = TRUE),stringsAsFactors=F)
  colnames(obsv2) <- c("phrases", "weight")
  obsv2$weight <- as.numeric(obsv2$weight)
  
  
  obsv3 <- merge(dfuniwords,obsv2, by="phrases",all.x = T)
  
  obsv3$weight[is.na(obsv3$weight)] <- 0
  
  obsv3$weight
}
```

```{r stmatrix}
dfuniwords <- data.frame(uniquewords,stringsAsFactors=F)
colnames(dfuniwords) <- c("phrases")
obsv <- str_to_lower(songstags$tags)
obsv <- str_split(obsv,pattern = ";")      #list

#tags-songs matrix, value is weight
obsv8 <- sapply(obsv,strtrans)     

#transpose to songs-tags matrix
stmatrix <- t(obsv8)             
dimnames(stmatrix) <- list(songstags$track_id,sort(uniquewords))
knitr::kable(
  head(stmatrix),
  caption = "songs-tags matrice"
)
```

The matrix size is `r nrow(stmatrix)`*`r ncol(stmatrix)`   

##Calculate songs-emotion matrix
```{r sematrix}
sematrix <- stmatrix %*% etmatrix1  
knitr::kable(
  head(sematrix),
  caption = "songs-emotion matrix"
)
```

The matrix size is `r nrow(sematrix)`*`r ncol(sematrix)`   


##Optimize sematrix data
* Fiter songs with tag value being zero    
 For some songs, the weight of emotion tag is 0. That's why some items in sematrix are 0-0-0-0.Use below method, we will remove these all-0 items. That is, the songs with weight zero are filtered out.
* Normalize data.  
 Due to big weight value, it is not good for reading and comparing.

```{r optisematrix}
sedf <- as.data.frame(sematrix)
#sedf$valid <- sedf$happy|sedf$sad|sedf$angry|sedf$calm
sedf$valid <- sedf$happy|sedf$sad|sedf$angry|sedf$relax
#nrow(sedf[!sedf$invalid,])  
sedf1 <- sedf[sedf$valid,]
sedf1$valid <- c()
sedf1 <- sedf1/100000L


sedfmax <- sapply(sedf1,max)
maxweight <- max(sedfmax)
sedfnorm <- sedf1/maxweight
sedfnorm1 <- sedfnorm*100
knitr::kable(
  head(sedfnorm1),
  caption = "Optimized songs-emotion dataset"
)
```

Now the total sample size is `r nrow(sedfnorm1)`   


```{r dimvalue}
N <- ncol(sedfnorm1)
avalue <- apply(sedfnorm1[,c('happy','sad','angry','relax')],1,function(x){(x[1]-x[2]+x[3]-x[4])/N/sqrt(2)})
vvalue <- apply(sedfnorm1[,c('happy','sad','angry','relax')],1,function(x){(x[1]-x[2]-x[3]+x[4])/N/sqrt(2)})
#avalue <- apply(sedfnorm1[,c('happy','sad','angry','calm')],1,function(x){(x[1]-x[2]+x[3]-x[4])/N/sqrt(2)})
#vvalue <- apply(sedfnorm1[,c('happy','sad','angry','calm')],1,function(x){(x[1]-x[2]-x[3]+x[4])/N/sqrt(2)})

adf <- as.data.frame(avalue)
adf <- cbind(track_id = rownames(adf), adf) 
rownames(adf) <- c()
adf$alevel <- ifelse(adf$avalue>=0,1,3)
vdf <- as.data.frame(vvalue)
vdf <- cbind(track_id = rownames(vdf), vdf) 
rownames(vdf) <- c()
vdf$vlevel <- ifelse(vdf$vvalue>=0,0,1)

dimvalue <- merge(vdf,adf,by='track_id')
dimvalue$quadrant <- dimvalue$vlevel+dimvalue$alevel
dimvalue$quadrant <- as.character(dimvalue$quadrant)
```
Songs emotion dataset based on 2D valence-arousal sapce  
```{r}
knitr::kable(
  head(dimvalue),
  caption = "2D songs-emotion dataset"
)
```

## Data Visualization
### macro view
```{r visualization_macro}
#the whole view
dimvalue%>%
  group_by(quadrant)%>%
  ggplot(aes(x = vvalue,y= avalue,color=quadrant))+
  geom_point(alpha=0.2)+
  scale_color_manual(values= c('orange','red','green','blue'),
                       name="Emotion Type",
                      breaks=c('1','2', '3','4'),
                      #labels=c("Happy", "Angry", "Calm","Sad")
                      labels=c("Happy", "Angry", "Relaxed","Sad")
                      )
```

### micro view
```{r visualization_micro}
dimvalue%>%
  group_by(quadrant)%>%
  ggplot(aes(x = vvalue,y= avalue,color=quadrant))+
  geom_point(alpha=0.2)+
  scale_color_manual(values= c('orange','red','green','blue'),
                     name="Emotion Type",
                     breaks=c('1','2', '3','4'),
                     #labels=c("Happy", "Angry", "Calm","Sad")
                     labels=c("Happy", "Angry", "Relaxed","Sad")
  )+
  coord_cartesian(xlim = c(-0.5,0.5),ylim = c(-0.5, 0.5))
```


## Statistical data overview for each quadrant
```{r statsummary}
summary_stats<- dimvalue%>%
  group_by(quadrant)%>%
  summarise(size=n())

knitr::kable(
  summary_stats
)
```

Note: 1 for happy, 2 for angry, 4 for sad, 3 for relax  

